Sampling requirements to detect ecosystem change

### 2020-03-19 R/manuscript_draft.Rmd
reorganizing and prepping for the discussion - I just couldn't focus on writing today

In line with theory about monitoring, we found that the probability of correcting identifying a changepoint decreased with smaller levels of population variability ($\sigma$) and autocorrelation ($\phi$). We also found that the probability of correct changepoint detection increases with larger shift sizes, which is essentially the effect size. Thus, the number of sample required to obtain high statistical power (above 0.8) increased with larger population variability, lower autocorrelation, and smaller shift sizes (Fig. \ref{fig:simulation_min_time}).

@@

Multiple subsampling strategies can be chosen(Fig. \ref{fig:conceptual}), but only iterative sampling allowed to approach the real changepoint in a limited number of samples (Fig. \ref{fig:simulation_min_time}c). Analyzing 11% of the sample was enough in most cases to approach the "true" changepoint.
In the context of a real case study, the iterative method allowed to find the main changepoint with 9 samples analyzed (\ref{fig:case_study_number_sample}), with a starting point of 5 randomly distributed samples. Same results for the major change, with only 12% of the samples analyzed. The method also work to detect further shifts (Supplementary Materials 1).
@Bruel2018 processed one sample at each centimeter in a 74-centimeter sediment core. Each sample took an average of 3 hours to process. We found that using an interval approach would have eliminated `r 3*(74-9)` hours of sample processing, or about `r round(3*(74-9)/8)` days, which is just a little over a month of work. This correspond to several thousands of US dollars depending on labor costs.

@@

Note that the number of sample needed varies depending on the characteristics of the time series. For perspective, the autocorrelation at lag-1 in the first DCA component we used in the case study example is `r round(ar_vr$acf[2],3)`, the variance is `r round(sd_vr,3)`, and the normalized amplitude of the shift was `r round(shift_vr,3)`.

<!--Limitations paragraph-->

@@

>>>>

In line with theory about monitoring, we found that the probability of correcting identifying a changepoint decreased with smaller levels of population variability ($\sigma$) and autocorrelation ($\phi$). We also found that the probability of correct changepoint detection increases with larger shift sizes, which is essentially the effect size. Thus, the number of sample required to obtain high statistical power (above 0.8) increased with larger population variability, lower autocorrelation, and smaller shift sizes (Fig. \ref{fig:simulation_min_time}). For perspective, the autocorrelation at lag-1 in the first DCA component we used in the case study example later is `r round(ar_vr$acf[2],3)`, the variance is `r round(sd_vr,3)`, and the normalized amplitude of the shift was `r round(shift_vr,3)`.

@@

<!---- reexplain and compare the 3 methods (theoretical (Fig 3) vs case study (Fig 4), and how they relate (Fig 2)) ---->
There is often a difference between the number of samples analyzed and the total number of samples collected, due to time or funding limitations. When the processing time is disproportionally higher than the collection time especially, a subsampling can be done posterior to sampling. A decision must then be taken of which subsamples to analyze.
We tested the performance at detecting changepoint of three subsampling methods: subsampling at regular and random interval, and an iterative sampling approach (Fig. \ref{fig:conceptual}).

We found that the iterative methods was systematically better at detecting changes than the two other methods, the random subsampling being the least efficient (Figs. \ref{simulation_number_sample} and \ref{fig:case_study_number_sample}). Autocorrelation, variance, and shift size, had an effect on how many samples were needed to detect the shift, but did not changed this conclusion.  (Fig. \ref{fig:simulation_min_time}). \rosalie{I'm curious to see, once you'll run Fig. 2 for the  3 methods, whether autocorrelation, shift size and variance really impact iterative method. It seems pretty robust!}

<!---- money implications ---->

Multiple subsampling strategies can be chosen(Fig. \ref{fig:conceptual}), but only iterative sampling allowed to approach the real changepoint in a limited number of samples (Fig. \ref{fig:simulation_min_time}c). Analyzing 11% of the sample was enough in most cases to approach the "true" changepoint.
In the context of a real case study, the iterative method allowed to find the main changepoint with 9 samples analyzed (Fig. \ref{fig:case_study_number_sample}), with a starting point of 5 randomly distributed samples. Same results for the major change, with only 12% of the samples analyzed. The method also work to detect further shifts (Supplementary Materials 1).
@Bruel2018 processed one sample at each centimeter in a 74-centimeter sediment core. Each sample took an average of 3 hours to process. We found that using an interval approach would have eliminated `r 3*(74-9)` hours of sample processing, or about `r round(3*(74-9)/8)` days, which is just a little over a month of work. This correspond to several thousands of US dollars depending on labor costs.

<!---- how this fits in with other work: examples of situation it would be applied (potential) ---->

@@

<!---- how this doesn't fit with other work (limitations) ---->

@@

<!---- conclusion ---->
Eventually, it is the researchers choice to choose their approach. Our objective was to quantify the implications of an approach vs. another. In the eventuality that the main objective is to detect a changepoint, we find that an approach is largely superior to the two others.
